---
title: "Cleaning Husking Satation Data"
author: "Camila Vargas"
date: "11/17/2021"
output: html_document
---

This scripts cleans, organize and QA/QC Alex Wegmann data on Husking Station Analysis data.

The outcome data of this script will be publicly available in the Environmental Data Initiative data portal.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(data.table)
library(tidyverse)
library(readxl)
library(janitor)
library(lubridate)
library(freeR)

## Sets R not to use scientific notations
options(scipen=999) 

```


## Read data
```{r}
excel_sheets("6.raw_data/HuskingStationAnalysis.xls")

rat_stations_raw <- read_excel(here::here("6.raw_data/HuskingStationAnalysis.xls"), sheet = "Rat_stations") %>%
  clean_names()

## Not needed see Alex respond below
# rat_crab_stat_raw <-  read_excel(here::here("6.raw_data/HuskingStationAnalysis.xls"), sheet = "Rat-crab stations") %>%
#   clean_names() %>% 
#   select(1:17)

```


We are only using data from 1 sheets: Rat Station. The rest seems to pull data form this two main data sets for a specific analysis.

Ask Alex: Do we need any other sheet? Is the data for ch 1 fig 3 sheet necessary or a compilation/restructure version of the two main sheets we are looking at.

Alex Response:
The “Rat_stations” tab is “parent dataset.” The Rat-crab stations tab is subset of stations where rats processed crab prey.  So, we only need to publish the Rat_stations dataset.

## Cleaning
Steps:
1. Change n/a to actual NAs
2. Remove units from colnames


```{r}

rat_station_clean <- rat_stations_raw %>% 
  ##remove number in front of variable name
  rename_at(vars(matches("number_")), ~str_remove(., "number_")) %>%
  rename_at(vars(12,15), ~str_remove(., "of_")) %>% 
  rename_at(vars(matches("_mm")), ~str_remove(., "_mm")) %>%
  rename(carapace_pieces = "carapace_pieces_y_n") %>%
  mutate(across(8:18, ~replace(.,.== "n/a", NA))) %>% 
  mutate(max_chela_size = ifelse(max_chela_size == "23,2", "23.2", max_chela_size) %>% as.numeric(.),
         ## remove all special characters
         contents = str_remove_all(contents, "[[:punct:]]"),
         total = ifelse(total == "PAF", NA, total) %>%  as.numeric(.),
         carapace_pieces = case_when(carapace_pieces == "Y" ~ "y",
                                     carapace_pieces == "N" ~ "n",
                                     T ~ carapace_pieces))

# rat_crab_clean <- rat_crab_stat_raw %>% 
#   rename_at(vars(matches("_mm")), ~str_remove(., "_mm")) %>% 
#   rename(carapace_pieces = "carapace_pieces_inc_gastrod_shell_used_by_coenobita_y_n") %>% 
#   mutate(across(7:17, ~replace(.,.== "n/a", NA))) %>% 
#   mutate(max_chela_size = ifelse(max_chela_size == "23,2", "23.2", max_chela_size),
#          contents = str_remove_all(contents, "[[:punct:]]"),
#          crab_genus = ifelse(crab_genus == "?", NA, crab_genus)) 

```


## Checking

```{r}
## Rat Husking
## counting how many NAs en each row. 
freeR::complete(rat_station_clean)

unique(rat_station_clean$island) ## Character
unique(rat_station_clean$forest) ## Categorical
unique(rat_station_clean$predator) ## Character
unique(rat_station_clean$prey) ## Character
unique(rat_station_clean$station) ## Numbers assign to each station
unique(rat_station_clean$contents) ## Has "" the need to be removed. Character
unique(rat_station_clean$total) ## What does PAF mean? otherwise numerical
unique(rat_station_clean$killed) ## numeric
unique(rat_station_clean$viable) ## read as numeric from saved file. same than seedling, chela and min_chela, gastropod_shells
unique(rat_station_clean$max_chela_size) ## one numbers had a , instead of a . fixed above.
unique(rat_station_clean$carapace_pieces) ## categorical
unique(rat_station_clean$station_habitat) ## categorical
unique(rat_station_clean$station_habitat_redifined) ## character. Except for "T


```


```{r}

## Rat-crab husking
## counting how many NAs en each row. 
# freeR::complete(rat_crab_clean)
# 
# unique(rat_crab_clean$island) ## Character
# unique(rat_crab_clean$forest) ## Categorical
# unique(rat_crab_clean$predator) ## Character
# unique(rat_crab_clean$prey) ## Character
# unique(rat_crab_clean$station) ## Numbers assign to each station. AKA interger
# unique(rat_crab_clean$contents) ## Character. Remove "" above
# unique(rat_crab_clean$crab_genus)## character. Replaced ? with NA
# unique(rat_crab_clean$total) ##  numeric
# unique(rat_crab_clean$killed) ## numeric
# unique(rat_crab_clean$chela) ## will be read as numeric
# unique(rat_crab_clean$max_chela_size) ## one numbers had a , instead of a . fixed above.
# unique(rat_crab_clean$carapace_pieces) ## will be read as numeric but should be categorical
# unique(rat_crab_clean$station_habitat_redifined) ## character.

```



## Save


```{r}
write_csv(rat_station_clean, here::here("7.clean_data/2004_2005_palmyra_atoll_rat_husking.csv"))

# write_csv(rat_crab_clean, here::here("7.clean_data/2004_2005_palmyra_atoll_rat_crab_husking.csv"))

```





